A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2016-11-15 EvoPar2017/paper.tex
Reviewed framework description

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for information 
  encoding, in this case chromosomes and related
  information sent by the clients and responses returned by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a
  key-value cache that uses as key the chromosome string itself and is
  reset when the solution is found. This cache has a finite 
  size that erases the oldest chromosomes once it has been filled to % Also "it has been filled to its"
  its maximum capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. 
  This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness chromosome could only be
  inserted once into the cache. That means that if a particular IP
  sends the fitness 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
\end{enumerate}

@@

NodIO needs a fitness function to work with; in this case the classical Trap 
function \cite{Ackley1987} has been used. 
The evolutionary algorithm that uses Trap as fitness function is
completely run in the
browser, but the server needs to know when the solution has been
found, so a subroutine that checks if the solution been found  
is also included in the server, as a configuration
option. In all cases we are using 50 Traps, a problem
that is difficult enough to need the intervention of several users for
an extended amount of minutes.
For every new individual sent by the clients to the cache, several
pieces of metadata are stored: a time stamp, the client IP, the chromosome and
fitness value, the cache size in that particular moment, and also if that
individual actually updated the cache or not.

All experiments are announced in the same way: we use social networks,
mainly Twitter, to announce the start of a new experiment. LinkedIn,
Facebook, and also private groups in Telegram and WhatsApp have been
used for announcements too. We also encourage friends and followers to
put an original post for announcing it, instead of just sharing, in
the way allowed by the platform, the original post. During the running
of the experiments, users are engaged, questions answered, and all
kind of explanations given if requested. In general, announcements
have been made in Spanish, but in some cases English was used
too. This makes for a certainly unreliable experimentation framework,
but it is a realistic one. The steps in the design of the evolutionary
algorithm will be explained next.

>>>>

The architecture of the {\sf NodIO} system is divided in two tiers:
\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for information 
  encoding, in this case chromosomes and related
  information sent by the clients, and responses returned by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a
  key-value cache that uses as key the chromosome string itself and is
  reset when the solution is found. This cache has a finite 
  size that erases the oldest chromosomes once it has been filled to % Also "it has been filled to its"
  its maximum capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. 
  This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness chromosome could only be
  inserted once into the cache. That means that if a particular IP
  sends a fitness value of 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This however did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
\end{enumerate}

@@

NodIO needs a fitness function to work with, in this case the classical Trap 
function \cite{Ackley1987} has been used. 
The evolutionary algorithm that uses Trap as fitness function is
completely run in the
browser, but the server also includes, as a configuration
option, a subroutine that checks if the solution has been
found. In all cases we are using 50 Traps, a problem
that is difficult enough to need the intervention of several users for
an extended amount of minutes. % Why? Do we have results on a single computer or on a cluster to compare?
For every new individual the clients send to the cache, several
pieces of metadata are stored: a time stamp, the client IP, the chromosome and
fitness value, the cache size in that particular moment, and also if that
individual actually updated the cache or not.

All experiments are announced in the same way: we use social networks,
mainly Twitter, to announce the start of a new experiment. LinkedIn,
Facebook, and also private groups in Telegram and WhatsApp have been
used for announcements too. We also encourage friends and followers to
put a self-authored post for announcing it, instead of just sharing, in
the way allowed by the platform, our original posts. During the running
of the experiments, users are engaged, questions answered, and all
kind of explanations given if requested. In general, announcements
have been made in Spanish, but in some cases English was used
too. This makes for a certainly unreliable experimentation framework,
but it is a realistic one. The steps in the design of the evolutionary
algorithm will be explained next.
